---
title: "Get study effort"
author: "Dan Benesh"
date: "7/2/2019"
output: html_document
---

I want to study host specificity in complex life cycle parasites. A key confounder of host specificity is study effort. Here I produce metrics for study effort on every parasite genus in the life cycle database.

```{r setup, include=FALSE}
library(ggplot2)
library(dplyr)
library(tidyr)
library(rentrez)
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
options(stringsAsFactors = FALSE)
theme_set(new = theme_bw())
```

```{r importdata}
dat_host <- read.csv(file = "../../data/CLC_database_hosts.csv", header = TRUE)
dat_lh <- read.csv(file = "../../data/CLC_database_lifehistory.csv", header = TRUE)
```

# Number of unique publications in life cycle database

```{r}
dat_host <- mutate(dat_host, publication = paste(Author, Year, Journal, Volume, Pages)) # creat publication variable
dat_lh <- mutate(dat_lh, publication = paste(Author, Year, Journal, Volume, Pages))
```
```{r}
dat_hosts_pub <- filter(dat_host, !is.na(Author), !is.na(Year), !is.na(Journal))%>%
  select(Parasite.genus, publication)%>%
  distinct()
dat_lh_pub <- filter(dat_lh, !is.na(Author), !is.na(Year), !is.na(Journal))%>%
  select(Parasite.genus, publication)%>%
  distinct()
study_effort <- bind_rows(dat_hosts_pub, dat_lh_pub)%>%arrange(Parasite.genus, publication)%>%distinct()
```
```{r}
study_effort <- group_by(study_effort, Parasite.genus)%>%
  summarise(n_lcdb = n())
```

```{r}
study_effort
```


# Pubmed hits


```{r}
# entrez_dbs()
# entrez_db_searchable("pubmed")
```

```{r}
# make queries for pubmed from species names
query_gen_name <- unique(dat_host$Parasite.genus)
query_gen_name <- gsub(query_gen_name, pattern = " ", replacement = "+")

# make additional query, species name plus parasite group
query_gen_name_group <- select(dat_host, Parasite.genus, Parasite.group)%>%distinct()
query_gen_name_group <- mutate(query_gen_name_group,
                              p_g_wc = if_else(Parasite.group == 'nematode', 'nematod*', 
                                               if_else(Parasite.group == 'cestode', 'cestod*', 'acanthocephala*')))
query_gen_name_group <- paste0(query_gen_name_group$Parasite.genus, '[ALL] AND ', query_gen_name_group$p_g_wc, '[ALL]')
```

First, I collect pubmed hits for just the genus name.

```{r}
for(query in query_gen_name){
  # search
  search_res <- entrez_search(db = "pubmed", term = query)
  # report hits
  num_hits <- search_res$count
  print(paste(num_hits, "hits for", query))
  
  if(query == query_gen_name[1]){
    pubmed_hits1 <- num_hits
  } else {
    pubmed_hits1 <- c(pubmed_hits1, num_hits)
  }
}

```

Then, I collect pubmed hits for the genus name + the taxonomic group.

```{r}
for(query in query_gen_name_group){
  # search
  search_res <- entrez_search(db = "pubmed", term = query)
  # report hits
  num_hits <- search_res$count
  print(paste(num_hits, "hits for", query))
  
  if(query == query_gen_name_group[1]){
    pubmed_hits2 <- num_hits
  } else {
    pubmed_hits2 <- c(pubmed_hits2, num_hits)
  }
}

```

```{r}
study_effort$n_pubmed_gen <- pubmed_hits1
study_effort$n_pubmed_gen_group <- pubmed_hits2
```

Here is the distribution of pubmed hits based on just genus names.

```{r}
# ggplot(study_effort, aes(x = n_lcdb)) + geom_histogram()
ggplot(study_effort, aes(x = n_pubmed_gen)) + geom_histogram()
```

The distribution based on genus name and parasite taxa looks similar.

```{r}
ggplot(study_effort, aes(x = n_pubmed_gen_group)) + geom_histogram()
```

Since it looks quite skewed, let's look at correlations on log scales. Here is the correlation between pubmed hits with and without a group term in the query. The correlation is quite good. At high values, the query including the group name returns fewer results (the black line is the 1:1 line). And there are a few genera where hits decrease substantially by adding a 'parasite group' term to the search.

```{r}
ggplot(study_effort, aes(y = n_pubmed_gen_group+1, x = n_pubmed_gen+1)) + 
  geom_point(alpha = 0.2) + 
  geom_smooth() +
  geom_abline(slope = 1, intercept = 0) +
  scale_y_log10() + scale_x_log10()
```

The genera with the largest relative difference between the two pubmed queries are shown. *Liga* stands out, because it is also a gene name.

```{r}
head( mutate(study_effort, diff_w_group_query = n_pubmed_gen - n_pubmed_gen_group,
             rel_diff_w_group_query = log10(n_pubmed_gen+1) - log10(n_pubmed_gen_group+1))%>%
        arrange(desc(rel_diff_w_group_query)), 15)
```

Are the number of pubmed hits correlated with the number of publications on a given genera in the life cycle database? It is, but it is noisy, which is not too surprising, as I was not trying to be exhaustive when compiling the database.

```{r}
ggplot(study_effort, aes(x = n_lcdb+1, y = n_pubmed_gen_group+1)) + 
  geom_point(alpha = 0.2) + 
  geom_smooth() +
  scale_y_log10() + scale_x_log10()
```

```{r}
write.csv(study_effort, file = "../../data/study_effort_genus.csv", row.names = F)
```

