---
title: "Look for tradeoffs driving host specificity"
output: github_document
---

Why are complex life cycle parasites specific at one stage but not another? Here I explore tradeoffs that might affect the evolution of host specificity in complex life cycles.

```{r, message=FALSE, warning=FALSE, include=FALSE}
library(dplyr)
library(ggplot2)
library(tidyr)
library(ape)
library(RColorBrewer)
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
options(stringsAsFactors = FALSE)
theme_set(new = theme_bw())
```

```{r importdata}
lcdb <- read.csv(file = "../data/CLC_database_updated_names.csv", header = TRUE)
st_level <- read.csv(file = "../data/stage_level_combined.csv", header = TRUE)
tree <- read.tree(file = "parasite_phylogeny/full_tree_time_calib.nex")
tree_tips <- read.csv(file = "../data/data_tree_tips_table.csv")
```

[Elsewhere](../stage_level_analyses/make_stage_level_df.Rmd), I created a data table at the level of stages within parasite species. I start by looking at correlations between variables measured at this level, like host records, taxonomic dissimilarity, development time, growth, etc. Here's a correlation matrix, but it might not be too useful, because several variables need log transformation.

```{r}
round(cor(select(st_level, study_effort, num_hosts_suspicious_removed, hsi_lcdb_suspcious_rem, avg_dd, rel_growth_len, rel_growth_biov), use = "pairwise"),2)
```

```{r}
tax.ranks <- c('genus', 'family', 'order', 'class', 'phylum') # for axis label
```

I calculated relative and absolute growth for parasite stages, as well as the time needed to complete development at a given stage, in days and degree days.

```{r}
# create starting sizes
st_level <- rename(st_level, stage_length_end = stage_body_length, stage_biov_end = stage_biov)%>%
  mutate(stage_length_start = stage_length_end - abs_growth_len,
         stage_biov_start = stage_biov_end - abs_growth_biov)
```

There is less development time data than growth data. And growth was calculated with two different size variables: length and biovolume. Biovolume had more missing data, because biovolume was calculated with lengths and widths and there were more missing widths than lengths.

```{r}
sapply(select(st_level, avg_dt, avg_dd, abs_growth_len, abs_growth_biov), function(x){sum(is.na(x))})
```

Therefore, for any paratenic host stage without growth data, I set the growth to zero and then replotted the data. This makes the negative relationship clearer, especially for worms in intermediate hosts (since those are the paratenic hosts).

```{r}
st_level <- mutate(st_level, rel_growth_paratenic = if_else(Facultative == "paratenic" & is.na(rel_growth_biov), 0, rel_growth_biov)) # if paratenic, then dt = 0
```

```{r}
st_level <- mutate(st_level, 
                   # if paratenic, then dt = 1 day at room temp
                   avg_dt_paratenic = if_else(Facultative == "paratenic" & is.na(avg_dt), 1, avg_dt),
                   avg_dd_paratenic = if_else(Facultative == "paratenic" & is.na(avg_dd), 20, avg_dd)) 
```

```{r}
st_level <- mutate(st_level, rel_growth_rate_dt = rel_growth_biov/avg_dt, rel_growth_rate_dd = rel_growth_biov/avg_dd) # make growth rate variable
```
```{r}
st_level <- mutate(st_level, Host_no_fac = if_else(Host.no > 4, as.integer(4), Host.no))%>%
  mutate(Host_no_fac = factor(Host_no_fac))
# center log-transformed study effort
st_level <- mutate(st_level, zstudy_effort = log10(study_effort+1) - mean( log10(study_effort+1), na.rm=T))
```
```{r}
st_level <- mutate(st_level, paratenic = if_else(Facultative == "paratenic", "yes", "no"),
                   imputed_gr = if_else(rel_growth_paratenic == 0, "yes", "no"),
                   imputed_dt = if_else(avg_dt_paratenic == 1, "yes", "no"))
```
```{r}
st_level <- mutate(st_level,
                   log_dd = log10(avg_dd),
                   log_dd_p = log10(avg_dd_paratenic),
                   log_dt = log10(avg_dt),
                   log_dt_p = log10(avg_dt_paratenic))
st_level$obs <- factor(1:length(st_level$Parasite.species)) # observation level effect for quantifying overdispersion
```


# Models

```{r}
library(lme4)
```

I want to test whether life history variables can explain variation in generalism, which would be evidence for a tradeoff. There are several groups of variables to test in the model: (1) uninteresting confounders (phylogeny, study effort), (2) life cycle variables (def vs int, host number), and (3) life history variables (growth and development). I first add the uninteresting confounders to the model, simply to control for them and to be sure that variables added afterwards explain *additional* variation. Here, I'm mainly interested in the life history variables. I could add them to the model either before or after the life cycle variables. If I add them after, I'm testing whether life history explains variation beyond that explained by the life cycle. If I add them before, I'm testing for life history tradeoffs, but their effects are possibly dependent/confounded with life cycle characteristics. I'll take both approaches below.

As for controlling for phylogeny, I added taxonomic categories into a linear mixed model. I did not include the highest and lower taxonomic levels (genus cuts the data quite thin while phylum is too coarse) as this improved model convergence. I still need to look at properly modelling phylogenetic effects

```{r}
r2_lmm_tax <- function(model) {
  # take in compound poisson mixed model, return marginal and conditional R2, ala Nakagawa et al. 2017
  # marginal r2 is just fixed effects
  # condition r2 is fixed and rand effects combined
  
  # model call
  call <- as.character(model@call)[2]
  
  # parameter estimates and df
  fixed_param <- fixef(model)
  df <- length(fixed_param) - 1
  
  # variance due to fixed effects
  pred <- as.vector(model.matrix(model) %*% fixed_param) # predicteds on basis of just fixed effects
  varF <- var(pred)
  
  
  # variance due to rand effects
  vc <- VarCorr(model)
  varSp <- vc$Parasite.species[1]
  if(length(vc) == 1){
    varR <- varSp
  } else {
    varR <- vc$Parasite.species[1] + vc$parasite_genus[1] + vc$parasite_family[1] + vc$parasite_order[1] + vc$parasite_class[1] + vc$parasite_phylum[1]
  }
  
  
  # residual var
  varE <- attr(vc, 'sc')^2
  
  
  # marginal r2
  mr2 <- varF/(varF + varR + varE)
  
  # conditional r2
  cr2 <- (varF + varR)/(varF + varR + varE)
  
  # conditional r2
  vexSP <- (varSp)/(varF + varR + varE)
  
  # output
  out_frame <- data_frame(call = call, df = df, marg_r2 = round(mr2, 3), cond_r2 = round(cr2,3), varSp = round(vexSP,3))
  return(out_frame)
}
```


# Taxonomic dissimilarity


Moving onto the next generalism variable: taxonomic dissimilarity.

### Case 1
#### Relative growth

As above, adding relative growth to a model with just taxonomy and study effort is an improvement. It has about the same magnitude effect as study effort. 
```{r}
filter(st_level, is.na(hsi_lcdb_suspcious_rem))
```

```{r}
reg00 <- lmer(hsi_lcdb_suspcious_rem ~ 1 + (1|Parasite.species),
             data = filter(st_level, !is.na(rel_growth_biov), !is.na(hsi_lcdb_suspcious_rem))) # just w/in species effect
reg0 <- update(reg00, . ~ . + (1|parasite_genus) + (1|parasite_family) + (1|parasite_order) + (1|parasite_class) + (1|parasite_phylum))
reg1 <- update(reg0, . ~ . + zstudy_effort) # add study effort (species level predictor)
reg2 <- update(reg1, . ~ . + rel_growth_biov) # add rel growth
reg3 <- update(reg2, . ~ . + poly(rel_growth_biov,2)) # add rel growth
anova(reg00, reg0, reg1, reg2, reg3)
```

Here are the parameters:

```{r}
summary(reg2)
```

```{r}
mod_list <- list(reg00, reg0, reg1, reg2, reg3)
if(exists("r2_table")){rm(r2_table)}
i <- 1
for(model in mod_list){
  if(i == 1){
    r2_table <- r2_lmm_tax(model)
  } else {
    r2_table <- rbind(r2_table, r2_lmm_tax(model))
  }
  i <- i + 1
}


r2_table <- mutate(r2_table, rand_var_explained = cond_r2 - marg_r2, df_used = df - lag(df))
r2_table$step <- c("within-species corr", "taxonomy", "study effort", "growth", "growth, quadratic")
r2_table <- select(r2_table, step, df_used, marg_r2, cond_r2, rand_var_explained, species_var_explained = varSp)
r2_table
```

The same pattern is seen if we assume worms do not grow in paratenic hosts, in fact the effect size is quite a bit larger.

```{r}
reg00p <- lmer(hsi_lcdb_suspcious_rem ~ 1 + (1|Parasite.species),
             data = filter(st_level, !is.na(rel_growth_paratenic), !is.na(hsi_lcdb_suspcious_rem))) # just w/in species effect
reg0p <- update(reg00p, . ~ . + (1|parasite_genus) + (1|parasite_family) + (1|parasite_order) + (1|parasite_class) + (1|parasite_phylum))
reg1p <- update(reg0p, . ~ . + zstudy_effort) # add study effort (species level predictor)
reg2p <- update(reg1p, . ~ . + rel_growth_paratenic) # add rel growth
reg3p <- update(reg1p, . ~ . + poly(rel_growth_paratenic,2)) # add rel growth
anova(reg00p, reg0p, reg1p, reg2p,reg3p)
```
```{r}
mod_list <- list(reg00p, reg0p, reg1p, reg2p, reg3p)
if(exists("r2_table")){rm(r2_table)}
i <- 1
for(model in mod_list){
  if(i == 1){
    r2_table <- r2_lmm_tax(model)
  } else {
    r2_table <- rbind(r2_table, r2_lmm_tax(model))
  }
  i <- i + 1
}


r2_table <- mutate(r2_table, rand_var_explained = cond_r2 - marg_r2, df_used = df - lag(df))
r2_table$step <- c("within-species corr", "taxonomy", "study effort", "growth", "growth, quadratic")
r2_table <- select(r2_table, step, df_used, marg_r2, cond_r2, rand_var_explained, species_var_explained = varSp)
r2_table
```

Plot results. Let's get best estimates and CIs from MCMCglmm

```{r}
datx <- filter(st_level, !is.na(rel_growth_biov), !is.na(hsi_lcdb_suspcious_rem))
datx$pred <- "no"
datx2 <- filter(st_level, !is.na(rel_growth_paratenic), !is.na(hsi_lcdb_suspcious_rem))
datx2$pred <- "no"
nd <- data.frame(zstudy_effort=0,
                 rel_growth_biov = seq(min(st_level$rel_growth_biov, na.rm = T),
                                            max(st_level$rel_growth_biov, na.rm = T),
                                            length.out = 50),
                 rel_growth_paratenic = seq(min(st_level$rel_growth_paratenic, na.rm = T),
                                            max(st_level$rel_growth_paratenic, na.rm = T),
                                            length.out = 50),
                 Parasite.species = datx$Parasite.species[1],
                 parasite_genus = datx$parasite_genus[1],
                 parasite_family = datx$parasite_family[1],
                 parasite_order = datx$parasite_order[1],
                 parasite_class = datx$parasite_class[1],
                 parasite_phylum = datx$parasite_phylum[1])
nd$pred <- "yes"
datx <- bind_rows(datx, nd)
datx2 <- bind_rows(datx2, nd)
```
```{r}
regp_mc1 <- MCMCglmm(hsi_lcdb_suspcious_rem ~ zstudy_effort + rel_growth_biov,
                    random = ~ Parasite.species + parasite_genus + parasite_family + 
                      parasite_order + parasite_class + parasite_phylum,
                    data = datx,
                    family = "gaussian",
                    nitt = 10500, thin = 30, burnin = 500) # just w/in species effect
regp_mc2 <- MCMCglmm(hsi_lcdb_suspcious_rem ~ zstudy_effort + rel_growth_paratenic,
                    random = ~ Parasite.species + parasite_genus + parasite_family + 
                      parasite_order + parasite_class + parasite_phylum,
                    data = datx2,
                    family = "gaussian",
                    nitt = 10500, thin = 30, burnin = 500) # just w/in species effect
```

Here are the fixed effects, `lmer` vs `MCMCglmm` for the model without imputation...

```{r}
cbind(fixef(reg2), summary(regp_mc1$Sol)$statistics[,1])
```

...and the model with imputation

```{r}
cbind(fixef(reg2p), summary(regp_mc2$Sol)$statistics[,1])
```

Here are the random effects `lmer` vs `MCMCglmm` for the model without imputation...

```{r}
vc <- as.data.frame(VarCorr(reg2))$vcov
cbind(round(vc,3), round(summary(regp_mc1$VCV)$statistics[,1],3))
rm(vc)
```

...and with imputation.

```{r}
vc <- as.data.frame(VarCorr(reg2p))$vcov
cbind(round(vc,3), round(summary(regp_mc2$VCV)$statistics[,1],3))
rm(vc)
```

Predict credible intervals

```{r}
p1 <- predict.MCMCglmm(regp_mc1,
                      type = "terms",
                      interval = "confidence",
                      marginal = ~ Parasite.species + parasite_genus + parasite_family + parasite_order +
                        parasite_class + parasite_phylum)
p1 <- bind_cols(datx, data.frame(p1))
p1 <- filter(p1, pred == "yes")
p1 <- mutate(p1, lwr = if_else(lwr < 1, 1, lwr))

p2 <- predict.MCMCglmm(regp_mc2,
                      type = "terms",
                      interval = "confidence",
                      marginal = ~ Parasite.species + parasite_genus + parasite_family + parasite_order +
                        parasite_class + parasite_phylum)
p2 <- bind_cols(datx2, data.frame(p2))
p2 <- filter(p2, pred == "yes")
p2 <- mutate(p2, lwr = if_else(lwr < 1, 1, lwr))
```


```{r}
# make model matrix
nd <- data.frame(zstudy_effort=0,
                 rel_growth_biov = seq(min(st_level$rel_growth_biov, na.rm = T),
                                       max(st_level$rel_growth_biov, na.rm = T),
                                       by = 0.2))
pd <- model.matrix(~zstudy_effort + rel_growth_biov, 
                   data = nd)

# bootstrap model
mySumm <- function(.) {
    c(beta=fixef(.))
}
bo <- bootMer(reg2, mySumm, nsim = 500) # bootstrap the model
bx <- as.data.frame(bo)
bx <- select(bx, starts_with('beta'))
bx <- as.matrix(bx)

boots <- bx %*% t(pd) # predicted values for every observation for every bootstrap
boots <- as.data.frame(t(boots))
names(boots) <- as.character(c(1:500))
boots <- bind_cols(nd, boots)
boots <- pivot_longer(boots, `1`:`500`)

# first get predicted size at each stage for every bootstrap
boot_means_fs <- boots%>%
  group_by(rel_growth_biov)%>%
  summarise(med_fs = mean(value), # mean and CI of the bootstraps
            high_fs = quantile(value, probs = 0.975),
            low_fs = quantile(value, probs = 0.025))
boot_means_fs <- mutate(boot_means_fs, low_fs = if_else(low_fs < 1, 1, low_fs))
```

```{r}
# make model matrix
nd <- data.frame(zstudy_effort=0,
                 rel_growth_paratenic = seq(min(st_level$rel_growth_paratenic, na.rm = T),
                                            max(st_level$rel_growth_paratenic, na.rm = T),
                                            by = 0.2))
pd <- model.matrix(~zstudy_effort + rel_growth_paratenic, 
                   data = nd)

# bootstrap model
bo <- bootMer(reg2p, mySumm, nsim = 500) # bootstrap the model
bx <- as.data.frame(bo)
bx <- select(bx, starts_with('beta'))
bx <- as.matrix(bx)

boots <- bx %*% t(pd) # predicted values for every observation for every bootstrap
boots <- as.data.frame(t(boots))
names(boots) <- as.character(c(1:500))
boots <- bind_cols(nd, boots)
boots <- pivot_longer(boots, `1`:`500`)

# first get predicted size at each stage for every bootstrap
boot_means_fs2 <- boots%>%
  group_by(rel_growth_paratenic)%>%
  summarise(med_fs = mean(value), # mean and CI of the bootstraps
            high_fs = quantile(value, probs = 0.975),
            low_fs = quantile(value, probs = 0.025))
boot_means_fs2 <- mutate(boot_means_fs2, low_fs = if_else(low_fs < 1, 1, low_fs))
```
```{r}
# # make model matrix
# nd <- data.frame(zstudy_effort=0,
#                  rel_growth_paratenic = seq(min(st_level$rel_growth_paratenic, na.rm = T),
#                                             max(st_level$rel_growth_paratenic, na.rm = T),
#                                             by = 0.2))
# pd <- model.matrix(~zstudy_effort + poly(rel_growth_paratenic,2), 
#                    data = nd)
# 
# # bootstrap model
# bo <- bootMer(reg3p, mySumm, nsim = 500) # bootstrap the model
# bx <- as.data.frame(bo)
# bx <- select(bx, starts_with('beta'))
# bx <- as.matrix(bx)
# 
# boots <- bx %*% t(pd) # predicted values for every observation for every bootstrap
# boots <- as.data.frame(t(boots))
# names(boots) <- as.character(c(1:500))
# boots <- bind_cols(nd, boots)
# boots <- pivot_longer(boots, `1`:`500`)
# 
# # first get predicted size at each stage for every bootstrap
# boot_means_fs3 <- boots%>%
#   group_by(rel_growth_paratenic)%>%
#   summarise(med_fs = mean(value), # mean and CI of the bootstraps
#             high_fs = quantile(value, probs = 0.975),
#             low_fs = quantile(value, probs = 0.025))
# boot_means_fs3 <- mutate(boot_means_fs2, low_fs = if_else(low_fs < 1, 1, low_fs))
```

Plots with predicted values. Predicted and CIs were basically the same with MCMC and LMER, either excluding imputed values...

```{r}
ggplot(filter(st_level, !is.na(rel_growth_paratenic)),
       aes(y = hsi_lcdb_suspcious_rem, x = rel_growth_paratenic)) +
  geom_point(aes(size = study_effort, shape = imputed_gr), 
             alpha = 0.1) +
  
  # PRED - MCMC WITHOUT IMP
  geom_ribbon(data = p1,
              aes(y = fit, x = rel_growth_biov,
                  ymin = lwr, ymax = upr),
              alpha = 0.2, fill = "red") +
  geom_line(data = p1, aes(y = fit, x = rel_growth_biov),
            linetype = "solid", color = "red") +
  
  # PRED - LMER WITHOUT IMP
  geom_ribbon(data = boot_means_fs,
              aes(y = med_fs, x = rel_growth_biov,
                  ymin = low_fs, ymax = high_fs),
              alpha = 0.5, fill = "gray") +
  geom_line(data = boot_means_fs, aes(y = med_fs, x = rel_growth_biov), 
            linetype = "solid", color = "black") +
  
  scale_shape_manual(values = c(16,17)) +
  scale_y_continuous(limits = c(1,6), breaks = c(1:6), labels = c("species", tax.ranks)) +
  theme(panel.grid.minor = element_blank()) +
  labs(x = "Relative growth\n(orders of magnitude size increase)", 
       shape = "Imputed?", y = "Taxonomic dissimilarity")
```

...or including them.

```{r}
ggplot(filter(st_level, !is.na(rel_growth_paratenic)),
       aes(y = hsi_lcdb_suspcious_rem, x = rel_growth_paratenic)) +
  geom_point(aes(size = study_effort, shape = imputed_gr), 
             alpha = 0.1) +
  
  # PRED - MCMC WITH IMP
  geom_ribbon(data = p2,
              aes(y = fit, x = rel_growth_paratenic,
                  ymin = lwr, ymax = upr),
              alpha = 0.2, fill = "red") +
  geom_line(data = p2, aes(y = fit, x = rel_growth_paratenic),
            linetype = "dashed", color = "red") +
  
  # PRED - LMER WITH IMP
  geom_ribbon(data = boot_means_fs2,
              aes(y = med_fs, x = rel_growth_paratenic,
                  ymin = low_fs, ymax = high_fs),
              alpha = 0.5, fill = "gray") +
  geom_line(data = boot_means_fs2, aes(y = med_fs, x = rel_growth_paratenic),
            linetype = "dashed", color = "black") +
  
  scale_shape_manual(values = c(16,17)) +
  scale_y_continuous(limits = c(1,6), breaks = c(1:6), labels = c("species", tax.ranks)) +
  theme(panel.grid.minor = element_blank()) +
  labs(x = "Relative growth\n(orders of magnitude size increase)", 
       shape = "Imputed?", y = "Taxonomic dissimilarity")
```

We'll stick with bootstrapped CIs from LMER as this is what the majority of paper is based on.

```{r}
f7a <- ggplot(filter(st_level, !is.na(rel_growth_paratenic)),
       aes(y = hsi_lcdb_suspcious_rem, x = rel_growth_paratenic)) +
  geom_point(aes(size = study_effort, shape = imputed_gr), 
             alpha = 0.1) +
  
  # PRED - LMER WITHOUT IMP
  geom_ribbon(data = boot_means_fs,
              aes(y = med_fs, x = rel_growth_biov,
                  ymin = low_fs, ymax = high_fs),
              alpha = 0.33, fill = "darkgray") +
  geom_line(data = boot_means_fs, aes(y = med_fs, x = rel_growth_biov), 
            linetype = "solid", color = "black") +
  
  # PRED - LMER WITH IMP
  geom_ribbon(data = boot_means_fs2,
              aes(y = med_fs, x = rel_growth_paratenic,
                  ymin = low_fs, ymax = high_fs),
              alpha = 0.33, fill = "gray") +
  geom_line(data = boot_means_fs2, aes(y = med_fs, x = rel_growth_paratenic),
            linetype = "dashed", color = "black") +
  # geom_ribbon(data = boot_means_fs3,
  #             aes(y = med_fs, x = rel_growth_paratenic,
  #                 ymin = low_fs, ymax = high_fs),
  #             alpha = 0.2) +
  # geom_line(data = boot_means_fs3, aes(y = med_fs, x = rel_growth_paratenic),
  #           linetype = "dotted") +
  guides(size = FALSE) +
  scale_shape_manual(values = c(16,17)) +
  scale_y_continuous(limits = c(1,6), breaks = c(1:6), labels = c("species", tax.ranks)) +
  theme(panel.grid.minor = element_blank()) +
  labs(x = "Relative growth\n(orders of magnitude size increase)", 
       shape = "Imputed?", y = "Taxonomic dissimilarity")
f7a
```


#### Development time - degree days

The taxonomic diversity of hosts is negatively related to developmental times (quick development - high generalism; long development - low generalism). 

```{r}
reg00d <- lmer(hsi_lcdb_suspcious_rem ~ 1 + (1|Parasite.species),
             data = filter(st_level, !is.na(avg_dd), !is.na(hsi_lcdb_suspcious_rem))) # just w/in species effect
reg0d <- update(reg00d, . ~ . + (1|parasite_genus) + (1|parasite_family) + (1|parasite_order) + (1|parasite_class) + (1|parasite_phylum))
reg1d <- update(reg0d, . ~ . + zstudy_effort) # add study effort (species level predictor)
reg2d <- update(reg1d, . ~ . + log_dd) # add devo
reg3d <- update(reg2d, . ~ . + I(log_dd^2)) # add devo
anova(reg00d, reg0d, reg1d, reg2d, reg3d)
```

Here are the parameters:
```{r}
summary(reg2d)
```

And the R^2^.

```{r}
mod_list <- list(reg00d, reg0d, reg1d, reg2d, reg3d)
if(exists("r2_table")){rm(r2_table)}
i <- 1
for(model in mod_list){
  if(i == 1){
    r2_table <- r2_lmm_tax(model)
  } else {
    r2_table <- rbind(r2_table, r2_lmm_tax(model))
  }
  i <- i + 1
}


r2_table <- mutate(r2_table, rand_var_explained = cond_r2 - marg_r2, df_used = df - lag(df))
r2_table$step <- c("within-species corr", "taxonomy", "study effort", "devo","devo,quad")
r2_table <- select(r2_table, step, df_used, marg_r2, cond_r2, rand_var_explained, species_var_explained = varSp)
r2_table
```

This is strongly exacerbated by assuming minimal development in paratenic hosts. Short development and generalism are associated.

```{r}
reg00dp <- lmer(hsi_lcdb_suspcious_rem ~ 1 + (1|Parasite.species),
             data = filter(st_level, !is.na(avg_dd_paratenic), !is.na(hsi_lcdb_suspcious_rem))) # just w/in species effect
reg0dp <- update(reg00dp, . ~ . + (1|parasite_genus) + (1|parasite_family) + (1|parasite_order) + (1|parasite_class) + (1|parasite_phylum))
reg1dp <- update(reg0dp, . ~ . + zstudy_effort) # add study effort (species level predictor)
reg2dp <- update(reg1dp, . ~ . + log_dd_p) # add devo
reg3dp <- update(reg2dp, . ~ . + I(log_dd_p^2)) # add devo
anova(reg00dp, reg0dp, reg1dp, reg2dp, reg3dp)
```
```{r}
mod_list <- list(reg00dp, reg0dp, reg1dp, reg2dp, reg3dp)
if(exists("r2_table")){rm(r2_table)}
i <- 1
for(model in mod_list){
  if(i == 1){
    r2_table <- r2_lmm_tax(model)
  } else {
    r2_table <- rbind(r2_table, r2_lmm_tax(model))
  }
  i <- i + 1
}


r2_table <- mutate(r2_table, rand_var_explained = cond_r2 - marg_r2, df_used = df - lag(df))
r2_table$step <- c("within-species corr", "taxonomy", "study effort", "devo", "devo,quad")
r2_table <- select(r2_table, step, df_used, marg_r2, cond_r2, rand_var_explained, species_var_explained = varSp)
r2_table
```

Look at same models, but MCMC.

```{r}
datx <- filter(st_level, !is.na(log_dd), !is.na(hsi_lcdb_suspcious_rem))
datx$pred <- "no"
datx2 <- filter(st_level, !is.na(log_dd_p), !is.na(hsi_lcdb_suspcious_rem))
datx2$pred <- "no"
nd <- data.frame(zstudy_effort=0,
                 log_dd = seq(min(st_level$log_dd, na.rm = T),
                              max(st_level$log_dd, na.rm = T),
                              length.out = 50),
                 log_dd_p = seq(min(st_level$log_dd_p, na.rm = T),
                                max(st_level$log_dd_p, na.rm = T),
                                length.out = 50),
                 Parasite.species = datx$Parasite.species[1],
                 parasite_genus = datx$parasite_genus[1],
                 parasite_family = datx$parasite_family[1],
                 parasite_order = datx$parasite_order[1],
                 parasite_class = datx$parasite_class[1],
                 parasite_phylum = datx$parasite_phylum[1])
nd$pred <- "yes"
datx <- bind_rows(datx, nd)
datx2 <- bind_rows(datx2, nd)
```
```{r}
regdt_mc1 <- MCMCglmm(hsi_lcdb_suspcious_rem ~ zstudy_effort + log_dd,
                    random = ~ Parasite.species + parasite_genus + parasite_family + 
                      parasite_order + parasite_class + parasite_phylum,
                    data = datx,
                    family = "gaussian",
                    nitt = 10500, thin = 30, burnin = 500) # just w/in species effect
regdt_mc2 <- MCMCglmm(hsi_lcdb_suspcious_rem ~ zstudy_effort + log_dd_p,
                    random = ~ Parasite.species + parasite_genus + parasite_family + 
                      parasite_order + parasite_class + parasite_phylum,
                    data = datx2,
                    family = "gaussian",
                    nitt = 10500, thin = 30, burnin = 500) # just w/in species effect
```

Here are the fixed effects, `lmer` vs `MCMCglmm` for the model without imputation...

```{r}
cbind(fixef(reg2d), summary(regdt_mc1$Sol)$statistics[,1])
```

...and the model with imputation

```{r}
cbind(fixef(reg2dp), summary(regdt_mc2$Sol)$statistics[,1])
```

Here are the random effects `lmer` vs `MCMCglmm` for the model without imputation...

```{r}
vc <- as.data.frame(VarCorr(reg2d))$vcov
cbind(round(vc,3), round(summary(regp_mc1$VCV)$statistics[,1],3))
rm(vc)
```

...and with imputation.

```{r}
vc <- as.data.frame(VarCorr(reg2dp))$vcov
cbind(round(vc,3), round(summary(regdt_mc2$VCV)$statistics[,1],3))
rm(vc)
```

Predict credible intervals

```{r}
p1 <- predict.MCMCglmm(regdt_mc1,
                      type = "terms",
                      interval = "confidence",
                      marginal = ~ Parasite.species + parasite_genus + parasite_family + parasite_order +
                        parasite_class + parasite_phylum)
p1 <- bind_cols(datx, data.frame(p1))
p1 <- filter(p1, pred == "yes")
p1 <- mutate(p1, lwr = if_else(lwr < 1, 1, lwr))

p2 <- predict.MCMCglmm(regdt_mc2,
                      type = "terms",
                      interval = "confidence",
                      marginal = ~ Parasite.species + parasite_genus + parasite_family + parasite_order +
                        parasite_class + parasite_phylum)
p2 <- bind_cols(datx2, data.frame(p2))
p2 <- filter(p2, pred == "yes")
p2 <- mutate(p2, lwr = if_else(lwr < 1, 1, lwr))
```


Now plots

```{r}
# make model matrix
nd <- data.frame(zstudy_effort=0,
                 log_dd = seq(min(st_level$log_dd, na.rm = T),
                              max(st_level$log_dd, na.rm = T),
                              length.out = 50))
pd <- model.matrix(~zstudy_effort + log_dd, 
                   data = nd)

# bootstrap model
mySumm <- function(.) {
    c(beta=fixef(.))
}
bo <- bootMer(reg2d, mySumm, nsim = 500) # bootstrap the model
bx <- as.data.frame(bo)
bx <- select(bx, starts_with('beta'))
bx <- as.matrix(bx)

boots <- bx %*% t(pd) # predicted values for every observation for every bootstrap
boots <- as.data.frame(t(boots))
names(boots) <- as.character(c(1:500))
boots <- bind_cols(nd, boots)
boots <- pivot_longer(boots, `1`:`500`)

# first get predicted size at each stage for every bootstrap
boot_means_dt <- boots%>%
  group_by(log_dd)%>%
  summarise(med_dt = mean(value), # mean and CI of the bootstraps
            high_dt = quantile(value, probs = 0.975),
            low_dt = quantile(value, probs = 0.025))
boot_means_dt <- mutate(boot_means_dt, low_dt = if_else(low_dt < 1, 1, low_dt))


# make model matrix
nd <- data.frame(zstudy_effort=0,
                 log_dd_p = seq(min(st_level$log_dd_p, na.rm = T),
                                max(st_level$log_dd_p, na.rm = T),
                                length.out = 50))
pd <- model.matrix(~zstudy_effort + log_dd_p, 
                   data = nd)

# bootstrap model
bo <- bootMer(reg2dp, mySumm, nsim = 500) # bootstrap the model
bx <- as.data.frame(bo)
bx <- select(bx, starts_with('beta'))
bx <- as.matrix(bx)

boots <- bx %*% t(pd) # predicted values for every observation for every bootstrap
boots <- as.data.frame(t(boots))
names(boots) <- as.character(c(1:500))
boots <- bind_cols(nd, boots)
boots <- pivot_longer(boots, `1`:`500`)

# first get predicted size at each stage for every bootstrap
boot_means_dt2 <- boots%>%
  group_by(log_dd_p)%>%
  summarise(med_dt = mean(value), # mean and CI of the bootstraps
            high_dt = quantile(value, probs = 0.975),
            low_dt = quantile(value, probs = 0.025))
boot_means_dt2 <- mutate(boot_means_dt2, low_dt = if_else(low_dt < 1, 1, low_dt))
```
```{r}
# # make model matrix
# nd <- data.frame(zstudy_effort=0,
#                  log_dd_p = seq(min(st_level$log_dd_p, na.rm = T),
#                                 max(st_level$log_dd_p, na.rm = T),
#                                 by = 0.2))
# pd <- model.matrix(~zstudy_effort + log_dd_p + I(log_dd_p^2), 
#                    data = nd)
# 
# # bootstrap model
# bo <- bootMer(reg3dp, mySumm, nsim = 500) # bootstrap the model
# bx <- as.data.frame(bo)
# bx <- select(bx, starts_with('beta'))
# bx <- as.matrix(bx)
# 
# boots <- bx %*% t(pd) # predicted values for every observation for every bootstrap
# boots <- as.data.frame(t(boots))
# names(boots) <- as.character(c(1:500))
# boots <- bind_cols(nd, boots)
# boots <- pivot_longer(boots, `1`:`500`)
# 
# # first get predicted size at each stage for every bootstrap
# boot_means_dt3 <- boots%>%
#   group_by(log_dd_p)%>%
#   summarise(med_dt = mean(value), # mean and CI of the bootstraps
#             high_dt = quantile(value, probs = 0.975),
#             low_dt = quantile(value, probs = 0.025))
# boot_means_dt3 <- mutate(boot_means_dt2, low_dt = if_else(low_dt < 1, 1, low_dt))
```

First, we check that MCMC and LMER gave the same CIs. Looks like it for unimputed...

```{r}
ggplot(filter(st_level, !is.na(log_dd_p), !is.na(hsi_lcdb_suspcious_rem)),
       aes(y = hsi_lcdb_suspcious_rem, x = 10^log_dd_p)) +
  geom_point(aes(size = study_effort, shape = imputed_dt),
             alpha = 0.1) +
  
  # PRED - MCMC WITHOUT IMP
  geom_ribbon(data = p1,
              aes(y = fit, x = 10^log_dd,
                  ymin = lwr, ymax = upr),
              alpha = 0.2, fill = "red") +
  geom_line(data = p1, aes(y = fit, x = 10^log_dd),
            linetype = "solid", color = "red") +
  
  # PRED - LMER WITHOUT IMP
  geom_ribbon(data = boot_means_dt,
              aes(y = med_dt, x = 10^log_dd,
                  ymin = low_dt, ymax = high_dt),
              alpha = 0.5, fill = "darkgray") +
  geom_line(data = boot_means_dt, aes(y = med_dt, x = 10^log_dd)) +
  
  scale_x_log10() +
  scale_shape_manual(values = c(16,17)) +
  scale_y_continuous(limits = c(1,6), breaks = c(1:6), labels = c("species", tax.ranks)) +
  theme(panel.grid.minor = element_blank()) +
  labs(x = "Development time (degree days)", 
       shape = "Imputed?", y = "Taxonomic dissimilarity")

```

...and imputed datasets.

```{r}
ggplot(filter(st_level, !is.na(log_dd_p), !is.na(hsi_lcdb_suspcious_rem)),
       aes(y = hsi_lcdb_suspcious_rem, x = 10^log_dd_p)) +
  geom_point(aes(size = study_effort, shape = imputed_dt),
             alpha = 0.1) +
  
  # PRED - MCMC WITH IMP
  geom_ribbon(data = p2,
              aes(y = fit, x = 10^log_dd_p,
                  ymin = lwr, ymax = upr),
              alpha = 0.2, fill = "red") +
  geom_line(data = p2, aes(y = fit, x = 10^log_dd_p),
            linetype = "dashed", color = "red") +

  # PRED - LMER WITH IMP
  geom_ribbon(data = boot_means_dt2,
              aes(y = med_dt, x = 10^log_dd_p,
                  ymin = low_dt, ymax = high_dt),
              alpha = 0.5, fill = "darkgray") +
  geom_line(data = boot_means_dt2, aes(y = med_dt, x = 10^log_dd_p),
            linetype = "dashed") +
  scale_x_log10() +
  scale_shape_manual(values = c(16,17)) +
  scale_y_continuous(limits = c(1,6), breaks = c(1:6), labels = c("species", tax.ranks)) +
  theme(panel.grid.minor = element_blank()) +
  labs(x = "Development time (degree days)", 
       shape = "Imputed?", y = "Taxonomic dissimilarity")
```

So let's just plot the LMER bootstrapped CIs.

```{r}
f7b <- ggplot(filter(st_level, !is.na(log_dd_p), !is.na(hsi_lcdb_suspcious_rem)),
       aes(y = hsi_lcdb_suspcious_rem, x = 10^log_dd_p)) +
  geom_point(aes(size = study_effort, shape = imputed_dt),
             alpha = 0.1) +
  
  # PRED - LMER WITHOUT IMP
  geom_ribbon(data = boot_means_dt,
              aes(y = med_dt, x = 10^log_dd,
                  ymin = low_dt, ymax = high_dt),
              alpha = 0.33, fill = "darkgray") +
  geom_line(data = boot_means_dt, aes(y = med_dt, x = 10^log_dd)) +
  
  # PRED - LMER WITH IMP
  geom_ribbon(data = boot_means_dt2,
              aes(y = med_dt, x = 10^log_dd_p,
                  ymin = low_dt, ymax = high_dt),
              alpha = 0.33, fill = "darkgray") +
  geom_line(data = boot_means_dt2, aes(y = med_dt, x = 10^log_dd_p),
            linetype = "dashed") +
  guides(size = F) +
  scale_x_log10() +
  scale_shape_manual(values = c(16,17)) +
  scale_y_continuous(limits = c(1,6), breaks = c(1:6), labels = c("species", tax.ranks)) +
  theme(panel.grid.minor = element_blank()) +
  labs(x = "Development time (degree days)", 
       shape = "Imputed?", y = "Taxonomic dissimilarity")
f7b
```


#### Development time - days

The taxonomic diversity of hosts is negatively related to developmental times (quick development - high generalism; long development - low generalism). 

```{r}
reg00d <- lmer(hsi_lcdb_suspcious_rem ~ 1 + (1|Parasite.species),
             data = filter(st_level, !is.na(avg_dt))) # just w/in species effect
reg0d <- update(reg00d, . ~ . + (1|parasite_genus) + (1|parasite_family) + (1|parasite_order) + (1|parasite_class) + (1|parasite_phylum))
reg1d <- update(reg0d, . ~ . + zstudy_effort) # add study effort (species level predictor)
reg2d <- update(reg1d, . ~ . + log_dt) # add devo
reg3d <- update(reg2d, . ~ . + I(log_dt^2)) # add devo
anova(reg00d, reg0d, reg1d, reg2d, reg3d)
```


And the R^2^.

```{r}
mod_list <- list(reg00d, reg0d, reg1d, reg2d, reg3d)
if(exists("r2_table")){rm(r2_table)}
i <- 1
for(model in mod_list){
  if(i == 1){
    r2_table <- r2_lmm_tax(model)
  } else {
    r2_table <- rbind(r2_table, r2_lmm_tax(model))
  }
  i <- i + 1
}


r2_table <- mutate(r2_table, rand_var_explained = cond_r2 - marg_r2, df_used = df - lag(df))
r2_table$step <- c("within-species corr", "taxonomy", "study effort", "devo","devo,quad")
r2_table <- select(r2_table, step, df_used, marg_r2, cond_r2, rand_var_explained, species_var_explained = varSp)
r2_table
```

This is strongly exacerbated by assuming minimal development in paratenic hosts. Short development and generalism are associated.

```{r}
reg00dp <- lmer(hsi_lcdb_suspcious_rem ~ 1 + (1|Parasite.species),
             data = filter(st_level, !is.na(avg_dt_paratenic))) # just w/in species effect
reg0dp <- update(reg00dp, . ~ . + (1|parasite_genus) + (1|parasite_family) + (1|parasite_order) + (1|parasite_class) + (1|parasite_phylum))
reg1dp <- update(reg0dp, . ~ . + zstudy_effort) # add study effort (species level predictor)
reg2dp <- update(reg1dp, . ~ . + log_dt_p) # add devo
reg3dp <- update(reg2dp, . ~ . + I(log_dt_p^2)) # add devo
anova(reg00dp, reg0dp, reg1dp, reg2dp, reg3dp)
```
```{r}
mod_list <- list(reg00dp, reg0dp, reg1dp, reg2dp, reg3dp)
if(exists("r2_table")){rm(r2_table)}
i <- 1
for(model in mod_list){
  if(i == 1){
    r2_table <- r2_lmm_tax(model)
  } else {
    r2_table <- rbind(r2_table, r2_lmm_tax(model))
  }
  i <- i + 1
}


r2_table <- mutate(r2_table, rand_var_explained = cond_r2 - marg_r2, df_used = df - lag(df))
r2_table$step <- c("within-species corr", "taxonomy", "study effort", "devo", "devo,quad")
r2_table <- select(r2_table, step, df_used, marg_r2, cond_r2, rand_var_explained, species_var_explained = varSp)
r2_table
```

Plots

```{r}
# make model matrix
nd <- data.frame(zstudy_effort=0,
                 log_dt = seq(min(st_level$log_dt, na.rm = T),
                              max(st_level$log_dt, na.rm = T),
                              by = 0.2))
pd <- model.matrix(~zstudy_effort + log_dt, 
                   data = nd)

# bootstrap model
mySumm <- function(.) { s <- sigma(.)
    c(beta =getME(., "beta"), sigma = s, sig01 = unname(s * getME(., "theta"))) }
bo <- bootMer(reg2d, mySumm, nsim = 500) # bootstrap the model
bx <- as.data.frame(bo)
bx <- select(bx, starts_with('beta'))
bx <- as.matrix(bx)

boots <- bx %*% t(pd) # predicted values for every observation for every bootstrap
boots <- as.data.frame(t(boots))
names(boots) <- as.character(c(1:500))
boots <- bind_cols(nd, boots)
boots <- pivot_longer(boots, `1`:`500`)

# first get predicted size at each stage for every bootstrap
boot_means_dt <- boots%>%
  group_by(log_dt)%>%
  summarise(med_dt = mean(value), # mean and CI of the bootstraps
            high_dt = quantile(value, probs = 0.975),
            low_dt = quantile(value, probs = 0.025))
boot_means_dt <- mutate(boot_means_dt, low_dt = if_else(low_dt < 1, 1, low_dt))
```

```{r}
# make model matrix
nd <- data.frame(zstudy_effort=0,
                 log_dt_p = seq(min(st_level$log_dt_p, na.rm = T),
                                max(st_level$log_dt_p, na.rm = T),
                                by = 0.2))
pd <- model.matrix(~zstudy_effort + log_dt_p, 
                   data = nd)

# bootstrap model
bo <- bootMer(reg2dp, mySumm, nsim = 500) # bootstrap the model
bx <- as.data.frame(bo)
bx <- select(bx, starts_with('beta'))
bx <- as.matrix(bx)

boots <- bx %*% t(pd) # predicted values for every observation for every bootstrap
boots <- as.data.frame(t(boots))
names(boots) <- as.character(c(1:500))
boots <- bind_cols(nd, boots)
boots <- pivot_longer(boots, `1`:`500`)

# first get predicted size at each stage for every bootstrap
boot_means_dt2 <- boots%>%
  group_by(log_dt_p)%>%
  summarise(med_dt = mean(value), # mean and CI of the bootstraps
            high_dt = quantile(value, probs = 0.975),
            low_dt = quantile(value, probs = 0.025))
boot_means_dt2 <- mutate(boot_means_dt2, low_dt = if_else(low_dt < 1, 1, low_dt))
```
```{r}
# # make model matrix
# nd <- data.frame(zstudy_effort=0,
#                  log_dt_p = seq(min(st_level$log_dt_p, na.rm = T),
#                                 max(st_level$log_dt_p, na.rm = T),
#                                 by = 0.2))
# pd <- model.matrix(~zstudy_effort + log_dt_p + I(log_dt_p^2), 
#                    data = nd)
# 
# # bootstrap model
# bo <- bootMer(reg3dp, mySumm, nsim = 500) # bootstrap the model
# bx <- as.data.frame(bo)
# bx <- select(bx, starts_with('beta'))
# bx <- as.matrix(bx)
# 
# boots <- bx %*% t(pd) # predicted values for every observation for every bootstrap
# boots <- as.data.frame(t(boots))
# names(boots) <- as.character(c(1:500))
# boots <- bind_cols(nd, boots)
# boots <- pivot_longer(boots, `1`:`500`)
# 
# # first get predicted size at each stage for every bootstrap
# boot_means_dt3 <- boots%>%
#   group_by(log_dt_p)%>%
#   summarise(med_dt = mean(value), # mean and CI of the bootstraps
#             high_dt = quantile(value, probs = 0.975),
#             low_dt = quantile(value, probs = 0.025))
# boot_means_dt3 <- mutate(boot_means_dt2, low_dt = if_else(low_dt < 1, 1, low_dt))
```
```{r}
ggplot(filter(st_level, !is.na(log_dt_p)),
       aes(y = hsi_lcdb_suspcious_rem, x = 10^log_dt_p)) +
  geom_point(aes(shape = paratenic), alpha = 0.1) +
  geom_ribbon(data = boot_means_dt,
              aes(y = med_dt, x = 10^log_dt,
                  ymin = low_dt, ymax = high_dt),
              alpha = 0.2) +
  geom_line(data = boot_means_dt, aes(y = med_dt, x = 10^log_dt)) +
  geom_ribbon(data = boot_means_dt2,
              aes(y = med_dt, x = 10^log_dt_p,
                  ymin = low_dt, ymax = high_dt),
              alpha = 0.2) +
  geom_line(data = boot_means_dt2, aes(y = med_dt, x = 10^log_dt_p),
            linetype = "dashed") +
  # geom_ribbon(data = boot_means_dt3,
  #             aes(y = med_dt, x = 10^log_dt_p,
  #                 ymin = low_dt, ymax = high_dt),
  #             alpha = 0.2, fill = "pink") +
  # geom_line(data = boot_means_dt3, aes(y = med_dt, x = 10^log_dt_p),
  #           linetype = "dotted", color = "pink") +
  guides(size = F) +
  scale_x_log10() +
  scale_shape_manual(values = c(16,17)) +
  scale_y_continuous(limits = c(1,6), breaks = c(1:6), labels = c("species", tax.ranks)) +
  theme(panel.grid.minor = element_blank()) +
  labs(x = "Development time (days)", 
       shape = "Paratenic host?", y = "Taxonomic dissimilarity")
```


#### Relative growth rate

```{r}
st_level <- mutate(st_level, 
                   # relative growth rate
                   rg = log(stage_biov_end) - log(stage_biov_start),
                   rg_dt = (log(stage_biov_end) - log(stage_biov_start))/avg_dt,
                   #rg_dt2 = (log10(stage_biov_end) - log10(stage_biov_start))/avg_dt, # perfectly correlates with rgr with ln
                   rg_dd = (log(stage_biov_end) - log(stage_biov_start))/avg_dd,
                   rg_rt = (log(stage_biov_end) - log(stage_biov_start))/(avg_dd/15))%>%
  # percent change in size per unit time
  mutate(rg_dt_perc = exp(rg_dt)-1, 
         rg_dd_perc = exp(rg_dd)-1,
         rg_rt_perc = exp(rg_rt)-1)%>% 
  # doubling times
  mutate(doubling_time_days = log(2)/rg_dt_perc,
         doubling_time_days_rt = log(2)/rg_rt_perc)
```

There are different ways of measuring growth rate. Here I use this formula for relative growth rate (rgr): (ln end size - ln starting size) / time. For modeling, it does not matter if we use log10 or ln, because the resulting rgrs are perfectly correlated. Interpretation of rgr is easier with natural logs, though, as it approximates the exponential growth rate (i.e. the % change per unit time), but only for lower growth rates. As growth accelerates, the relationship between rgr and the exponential growth rate (% change per day) diverges a bit.

```{r}
ggplot(st_level, 
       aes(x = rg_rt, y = rg_rt_perc),
       ) +
  geom_point( alpha = 0.3) +
  geom_abline() +
  labs(x = "Relative growth rate (per day at room temp)", y = "Compound growth (% change per day at room temp)")
```

We can also convert ln-based rgr to doubling time [ ln(2)/exponential growth rate ], which is perhaps more intuitive. However, since some worms decreased in size, doubling time is halving time in these cases. Here is the distribution of doubling times. It takes about 5 days at room temp to double in size. Most of the half-lives are long, given worms decreases only slightly in size.

```{r}
ggplot(st_level,
       aes(x = abs(doubling_time_days_rt))) +
  geom_histogram(aes(fill = doubling_time_days_rt<0)) +
  scale_x_log10(breaks = c(1, 3, 5, 7, 10, 25, 50, 100)) +
  scale_fill_discrete(labels = c("doubling time", "half-life")) +
  theme(panel.grid.minor = element_blank()) +
  labs(x = "Days at room temp", fill = NULL)
```

Fit the models. Growth rate was unrelated to taxonomic dissimilarity, explaining no variation.

```{r}
reg00rg <- lmer(hsi_lcdb_suspcious_rem ~ 1 + (1|Parasite.species),
             data = filter(st_level, !is.na(rg_dd), !is.na(hsi_lcdb_suspcious_rem))) # just w/in species effect
reg0rg <- update(reg00rg, . ~ . + (1|parasite_genus) + (1|parasite_family) + (1|parasite_order) + (1|parasite_class) + (1|parasite_phylum))
reg1rg <- update(reg0rg, . ~ . + zstudy_effort) # add study effort (species level predictor)
reg2rg <- update(reg1rg, . ~ . + rg_dd) # add rel growth
anova(reg00rg, reg0rg, reg1rg, reg2rg)
```

```{r}
mod_list <- list(reg00rg, reg0rg, reg1rg, reg2rg)
if(exists("r2_table")){rm(r2_table)}
i <- 1
for(model in mod_list){
  if(i == 1){
    r2_table <- r2_lmm_tax(model)
  } else {
    r2_table <- rbind(r2_table, r2_lmm_tax(model))
  }
  i <- i + 1
}


r2_table <- mutate(r2_table, rand_var_explained = cond_r2 - marg_r2, df_used = df - lag(df))
r2_table$step <- c("within-species corr", "taxonomy", "study effort", "rel growth rate")
r2_table <- select(r2_table, step, df_used, marg_r2, cond_r2, rand_var_explained, species_var_explained = varSp)
r2_table
```

It does not matter if we change the way we measure relative growth. Such as by using days instead of degree days.

```{r}
reg2rgx <- update(reg1rg, . ~ . + rg_dt) # add rel growth
anova(reg1rg, reg2rgx)
```

Or using the percent change in size per day at room temp.

```{r}
reg2rgx <- update(reg1rg, . ~ . + rg_rt_perc) # add rel growth
anova(reg1rg, reg2rgx)
```

Or using doubling time instead of rgr.

```{r}
reg2rgy <- update(reg1rg, . ~ . + doubling_time_days_rt) # add rel growth
anova(reg1rg, reg2rgy)
```

The same null result is seen if we assume worms do not grow in paratenic hosts. But this does not make that much sense, since calculating growth rates is suspect for stages that lack growth.

```{r}
st_level <- mutate(st_level, rel_growth_rate_paratenic = rel_growth_paratenic/avg_dd_paratenic) # make growth rate variable
```
```{r}
reg00rgp <- lmer(hsi_lcdb_suspcious_rem ~ 1 + (1|Parasite.species),
             data = filter(st_level, !is.na(rel_growth_rate_paratenic), !is.na(hsi_lcdb_suspcious_rem)))
reg0rgp <- update(reg00rgp, . ~ . + (1|parasite_genus) + (1|parasite_family) + (1|parasite_order) + (1|parasite_class) + (1|parasite_phylum))
reg1rgp <- update(reg0rgp, . ~ . + zstudy_effort) # add study effort (species level predictor)
reg2rgp <- update(reg1rgp, . ~ . + rel_growth_rate_paratenic) # add rel growth
anova(reg00rgp, reg0rgp, reg1rgp, reg2rgp)
```
```{r}
mod_list <- list(reg00rgp, reg0rgp, reg1rgp, reg2rgp)
if(exists("r2_table")){rm(r2_table)}
i <- 1
for(model in mod_list){
  if(i == 1){
    r2_table <- r2_lmm_tax(model)
  } else {
    r2_table <- rbind(r2_table, r2_lmm_tax(model))
  }
  i <- i + 1
}


r2_table <- mutate(r2_table, rand_var_explained = cond_r2 - marg_r2, df_used = df - lag(df))
r2_table$step <- c("within-species corr", "taxonomy", "study effort", "devo")
r2_table <- select(r2_table, step, df_used, marg_r2, cond_r2, rand_var_explained, species_var_explained = varSp)
r2_table
```


```{r}
# make model matrix
nd <- data.frame(zstudy_effort=0,
                 rg_dd = seq(min(st_level$rg_dd, na.rm = T),
                                          max(st_level$rg_dd, na.rm = T),
                                          by = 0.001))
pd <- model.matrix(~zstudy_effort + rg_dd, 
                   data = nd)

# bootstrap model
bo <- bootMer(reg2rg, mySumm, nsim = 500) # bootstrap the model
bx <- as.data.frame(bo)
bx <- select(bx, starts_with('beta'))
bx <- as.matrix(bx)

boots <- bx %*% t(pd) # predicted values for every observation for every bootstrap
boots <- as.data.frame(t(boots))
names(boots) <- as.character(c(1:500))
boots <- bind_cols(nd, boots)
boots <- pivot_longer(boots, `1`:`500`)

# first get predicted size at each stage for every bootstrap
boot_means_rg <- boots%>%
  group_by(rg_dd)%>%
  summarise(med_rg = mean(value), # mean and CI of the bootstraps
            high_rg = quantile(value, probs = 0.975),
            low_rg = quantile(value, probs = 0.025))
```
```{r}
# make model matrix
nd <- data.frame(zstudy_effort=0,
                 rel_growth_rate_paratenic = seq(min(st_level$rel_growth_rate_paratenic, na.rm = T),
                                                 max(st_level$rel_growth_rate_paratenic, na.rm = T),
                                                 by = 0.002))
pd <- model.matrix(~zstudy_effort + rel_growth_rate_paratenic, 
                   data = nd)

# bootstrap model
bo <- bootMer(reg2rgp, mySumm, nsim = 500) # bootstrap the model
bx <- as.data.frame(bo)
bx <- select(bx, starts_with('beta'))
bx <- as.matrix(bx)

boots <- bx %*% t(pd) # predicted values for every observation for every bootstrap
boots <- as.data.frame(t(boots))
names(boots) <- as.character(c(1:500))
boots <- bind_cols(nd, boots)
boots <- pivot_longer(boots, `1`:`500`)

# first get predicted size at each stage for every bootstrap
boot_means_rg2 <- boots%>%
  group_by(rel_growth_rate_paratenic)%>%
  summarise(med_rg = mean(value), # mean and CI of the bootstraps
            high_rg = quantile(value, probs = 0.975),
            low_rg = quantile(value, probs = 0.025))
```

Get some weird growth rates when we impute data for paratenic hosts.

```{r}
ggplot(filter(st_level, !is.na(rel_growth_rate_paratenic)),
       aes(y = hsi_lcdb_suspcious_rem, x = rel_growth_rate_paratenic)) +
  geom_point(aes(shape = paratenic),
             alpha = 0.1) +
  geom_ribbon(data = boot_means_rg,
              aes(y = med_rg, x = rg_dd,
                  ymin = low_rg, ymax = high_rg),
              alpha = 0.2) +
  geom_line(data = boot_means_rg, aes(y = med_rg, x = rg_dd)) +
  geom_ribbon(data = boot_means_rg2,
              aes(y = med_rg, x = rel_growth_rate_paratenic,
                  ymin = low_rg, ymax = high_rg),
              alpha = 0.2) +
  geom_line(data = boot_means_rg2, aes(y = med_rg, x = rel_growth_rate_paratenic),
            linetype = "dashed") +
  scale_shape_manual(values = c(16,17)) +
  scale_y_continuous(limits = c(1,6), breaks = c(1:6), labels = c("species", tax.ranks)) +
  theme(panel.grid.minor = element_blank()) +
  labs(x = "Relative growth rate", 
       shape = "Paratenic host?", y = "Taxonomic dissimilarity")
```

Let's therefore only plot the worms with clear growth. And to make the plot more interpretable, we'll change the x-axis from relative growth in degree days to relative growth in days at room temp (20).

```{r}
reg2rgx <- update(reg1rg, . ~ . + rg_rt) # add rel growth
```
```{r}
# make model matrix
nd <- data.frame(zstudy_effort=0,
                 rg_rt = seq(min(st_level$rg_rt, na.rm = T),
                             max(st_level$rg_rt, na.rm = T),
                             length.out = 50))
pd <- model.matrix(~zstudy_effort + rg_rt, 
                   data = nd)

# bootstrap model
bo <- bootMer(reg2rgx, mySumm, nsim = 500) # bootstrap the model
bx <- as.data.frame(bo)
bx <- select(bx, starts_with('beta'))
bx <- as.matrix(bx)

boots <- bx %*% t(pd) # predicted values for every observation for every bootstrap
boots <- as.data.frame(t(boots))
names(boots) <- as.character(c(1:500))
boots <- bind_cols(nd, boots)
boots <- pivot_longer(boots, `1`:`500`)

# first get predicted size at each stage for every bootstrap
boot_means_rg <- boots%>%
  group_by(rg_rt)%>%
  summarise(med_rg = mean(value), # mean and CI of the bootstraps
            high_rg = quantile(value, probs = 0.975),
            low_rg = quantile(value, probs = 0.025))
```


```{r}
f7cc <- ggplot(filter(st_level, !is.na(rg_rt)),
       aes(y = hsi_lcdb_suspcious_rem, x = rg_rt)) +
  geom_point(aes(size = study_effort), alpha = 0.1) +
  geom_ribbon(data = boot_means_rg,
              aes(y = med_rg, x = rg_rt,
                  ymin = low_rg, ymax = high_rg),
              alpha = 0.33, fill = "darkgray") +
  guides(size = F) +
  geom_line(data = boot_means_rg, aes(y = med_rg, x = rg_rt)) +
  scale_shape_manual(values = c(16)) +
  scale_y_continuous(limits = c(1,6), breaks = c(1:6), labels = c("species", tax.ranks)) +
  theme(panel.grid.minor = element_blank()) +
  labs(x = "Relative growth rate\n(~% size increase per day at 20C)", y = "Taxonomic dissimilarity")
f7cc
```

# Growth rate as combo of growth and devo

```{r}
st_level <- mutate(st_level, generalist_fac = cut(hsi_lcdb_suspcious_rem, breaks = c(0, 1, 2, 3, 4, 6)))%>% # cut tax dissim by tax hierarchy
  mutate(generalist_fac = factor(generalist_fac, labels = c("species", "genus", "family", "order", "class+")))
```
```{r}
st_level_dt <- group_by(st_level, generalist_fac)%>%
  summarize(avg_dd_p = mean(log10(avg_dd_paratenic), na.rm = T),
            sd_dd_p = sd(log10(avg_dd_paratenic), na.rm = T),
            n_dt = sum(!is.na(avg_dd_paratenic)),
            
            avg_g_p = mean(rel_growth_paratenic, na.rm = T),
            sd_g_p = sd(rel_growth_paratenic, na.rm = T),
            n_g = sum(!is.na(rel_growth_paratenic))
            )
st_level_dt <- mutate(st_level_dt, se_dd_p = sd_dd_p/sqrt(n_dt), se_g_p = sd_g_p/sqrt(n_g))%>%na.omit()
```

If parasite stages infect hosts from different classes, then they exhibit short development and little growth. By contrast, if hosts for a given parasite stage tend to be of the same genus or family, then the parasite stage tend to spend time growing.

```{r}
f7c <- ggplot(st_level, aes(x = avg_dd_paratenic, y = rel_growth_paratenic)) +
  scale_x_log10() + 
  geom_point(aes(shape = paratenic), alpha = 0.1) +
  geom_point(data = st_level_dt, aes(x = 10^(avg_dd_p), y = avg_g_p, color = generalist_fac),
             alpha = 0.75, size = 2.5) +
  geom_errorbar(data = st_level_dt, aes(x = 10^(avg_dd_p), y = avg_g_p, 
                                        ymin = avg_g_p - 1.96*se_g_p, ymax = avg_g_p + 1.96*se_g_p,
                                        color = generalist_fac),
                alpha = 0.75, size = 1, width = 0) +
  geom_errorbarh(data = st_level_dt, aes(x = 10^(avg_dd_p), y = avg_g_p,
                                         xmin = 10^(avg_dd_p - 1.96*se_dd_p), 
                                         xmax = 10^(avg_dd_p + 1.96*se_dd_p),
                                         color = generalist_fac), 
                alpha = 0.75, size = 1, height = 0) +
  # geom_smooth(method = lm, color = 'black', linetype = 'dashed', se = F) +
  scale_shape_manual(values = c(16,17)) +
  scale_color_brewer(type = 'qual', palette = "Dark2") +
  # scale_color_manual(values = mypal) +
  theme(panel.grid.minor = element_blank()) +
  labs(y = "Relative growth\n(orders of magnitude size increase)", x = "Development time (degree days)", 
       shape = "Paratenic host?", color = "Taxonomic\ndissimilarity") +
  guides(shape = guide_legend(override.aes = list(size = 4, alpha = 0.25))) +
  annotate(geom = 'text', label = '(c)', x = min(st_level$avg_dd_paratenic, na.rm=T), y = max(st_level$rel_growth_paratenic, na.rm=T) )
f7c
```

We can also take the residuals of the above plot (faster or slower growth than expected) and see if they are related to generalism. They are not.

```{r}
dx1 <- filter(st_level, !is.na(log_dd), !is.na(rel_growth_biov))
modx1 <- lm(rel_growth_biov ~ log_dd, data = dx1)
dx1$res1 <- residuals((modx1))

dx2 <- filter(st_level, !is.na(log_dd_p), !is.na(rel_growth_paratenic))
modx2 <- lm(rel_growth_paratenic ~ log_dd_p, data = dx2)
dx2$res2 <- residuals((modx2))
dx <- left_join(dx2, select(dx1, obs, res1))
```

```{r}
ffg <-ggplot(dx, aes(x = res2, y = hsi_lcdb_suspcious_rem)) +
  geom_point(alpha = 0.3, aes(shape = paratenic)) +
  geom_smooth(data = filter(dx, paratenic == "no"), method = lm, se = F) +
  geom_smooth(method = lm, se = F, linetype = "dashed") +
  scale_y_continuous(limits = c(1,6), breaks = c(1:6), labels = c("species", tax.ranks)) +
  theme(panel.grid.minor = element_blank()) +
  labs(x = "Residuals, Rel Growth vs Log DD", y = "Taxonomic dissimilarity")
ffg
# ggsave(ffg, filename = "../../figs/growth_rate_resid_plot.png")
```




```{r}
library(cowplot)
```

```{r}
f7a2 <- f7a + 
  guides(size = F, shape = F) +
  annotate(geom = 'text', label = '(a)', x = min(st_level$rel_growth_paratenic, na.rm=T), y = 6 )
f7b2 <- f7b + 
  guides(size = F) +
  annotate(geom = 'text', label = '(b)', x = min(st_level$avg_dd_paratenic, na.rm=T), y = 6 ) +
  theme(axis.title.y = element_blank(),
        axis.text.y = element_blank(),
        legend.background = element_rect(color = "black"),
        # legend.direction = "horizontal",
        legend.position = c(0.85, 0.85)) +
  guides(shape = guide_legend(override.aes = list(alpha = 1)))
f7cc2 <- f7cc +
  guides(size = F) +
  annotate(geom = 'text', label = '(c)', x = min(st_level$rg_rt, na.rm=T), y = 6 ) +
  theme(axis.title.y = element_blank(),
        axis.text.y = element_blank())
```
```{r}
f7_v1 <- plot_grid(f7a2, f7b2,  nrow = 1, align = "hv")
f7_v2 <- plot_grid(f7a2, f7b2, f7c, nrow = 1, align = "hv", rel_heights = c(1,1,1.5))
f7_v3 <- plot_grid(f7a2, f7b2, f7cc2, nrow = 1, align = "hv", axis = "tb")
```
```{r}
fig_width <- 4.5
fig_height <- 4.5

# ggsave(f7_v1, filename = "../../figs/fig7_rev_v1.png", width = fig_width*2, height = fig_height)
# ggsave(f7_v2, filename = "../../figs/fig7_rev_v2.png", width = fig_width*3, height = fig_height)
ggsave(f7_v3, filename = "../../figs/fig6_rev.png", width = fig_width*3, height = fig_height)
ggsave(f7_v3, filename = "../../figs/fig6_rev.svg", width = fig_width*3, height = fig_height)
# in svg move figs closer together, put A on top first, then B, then move
```


# Conclusions

Parasite life cycle, life history, and patterns of host generalism are related. In particular, parasite exhibit high generalism in second or third intermediate hosts. These tend to be paratenic hosts, in which parasites undergo little growth and development. But even if these hosts are excluded, there is still a negative relationship between generalism and worm growth. Worms grow less in stages where they infect a broader range of hosts, which is consistent with costs of generalism.