---
title: "Look for tradeoffs driving host specificity"
output: github_document
---

Why are complex life cycle parasites specific at one stage but not another? Here I explore tradeoffs that might affect the evolution of host specificity in complex life cycles.

```{r, message=FALSE, warning=FALSE, include=FALSE}
library(dplyr)
library(ggplot2)
library(tidyr)
library(ape)
library(RColorBrewer)
knitr::opts_chunk$set(echo = FALSE, warning = FALSE, message = FALSE)
options(stringsAsFactors = FALSE)
theme_set(new = theme_bw())
```

```{r importdata}
lcdb <- read.csv(file = "../../data/CLC_database_updated_names.csv", header = TRUE)
st_level <- read.csv(file = "../../data/stage_level_combined.csv", header = TRUE)
tree <- read.tree(file = "../parasite_phylogeny/full_tree_time_calib.nex")
tree_tips <- read.csv(file = "../../data/data_tree_tips_table.csv")
```

[Elsewhere](../stage_level_analyses/make_stage_level_df.Rmd), I created a data table at the level of stages within parasite species. I start by looking at correlations between variables measured at this level, like host records, taxonomic dissimilarity, development time, growth, etc. Here's a correlation matrix, but it might not be too useful, because several variables need log transformation.

```{r}
round(cor(select(st_level, study_effort, num_hosts_suspicious_removed, hsi_lcdb_suspcious_rem, avg_dd, rel_growth_len, rel_growth_biov), use = "pairwise"),2)
```

## Correlations among specificity variables

The host generalism measures are calculated with only the life cycle database, because stage info is not present in the NHM data. The two measures of generalism, number of host species and taxonomic dissimilarity of host species, are positively correlated; stages with more host species (host range) have been recorded from more taxonomically diverse hosts.

```{r}
tax.ranks <- c('genus', 'family', 'order', 'class', 'phylum') # for axis label
```
```{r}
ggplot(st_level, aes(x = num_hosts_suspicious_removed, y = hsi_lcdb_suspcious_rem)) +
  scale_x_log10() +
  scale_y_continuous(limits = c(1,6), breaks = c(1:6), labels = c("species", tax.ranks)) +
  geom_point(alpha = 0.2) +
  geom_smooth() +
  labs(x = "Host records", y = "Tax dissimilarity index", title = "Stage level")
```

The pattern holds if we exclude stages recorded from just 1 host species. Thus, the two measures are not independent.

```{r}
ggplot(filter(st_level, num_hosts_suspicious_removed > 1),
       aes(x = num_hosts_suspicious_removed, y = hsi_lcdb_suspcious_rem)) +
  scale_x_log10() +
  scale_y_continuous(limits = c(1,6), breaks = c(1:6), labels = c("species", tax.ranks)) +
  geom_point(alpha = 0.2) +
  geom_smooth() +
  labs(x = "Host records", y = "Tax dissimilarity index", title = "Stage level, n = 1 host removed")
```

Study effort is still an important confounder of generalism measures - the host range increases with study effort, perhaps more for intermediate hosts, than definitive hosts. That sorta makes sense, since fewer studies are on intermediate hosts, it could take more to add new host records.

```{r}
ggplot(st_level, aes(y = num_hosts_suspicious_removed, x = study_effort+1, color = Def.int)) +
  scale_x_log10() + scale_y_log10() +
  geom_point(alpha = 0.2) +
  geom_smooth() +
  labs(x = "Study effort", y = "Host records", title = "Stage level")
```

Study effort also increases the diversity of hosts at a given stage.

```{r}
ggplot(st_level, aes(y = hsi_lcdb_suspcious_rem, x = study_effort+1, color = Def.int)) +
  scale_x_log10() + scale_y_log10() +
  geom_point(alpha = 0.2) +
  geom_smooth() +
  labs(x = "Study effort", y = "Tax dissimilarity index", title = "Stage level")
```


## Correlations among life history variables

I calculated relative and absolute growth for parasite stages, as well as the time needed to complete development at a given stage, in days and degree days.

```{r}
# create starting sizes
st_level <- rename(st_level, stage_length_end = stage_body_length, stage_biov_end = stage_biov)%>%
  mutate(stage_length_start = stage_length_end - abs_growth_len,
         stage_biov_start = stage_biov_end - abs_growth_biov)
```

There is less development time data than growth data. And growth was calculated with two different size variables: length and biovolume. Biovolume had more missing data, because biovolume was calculated with lengths and widths and there were more missing widths than lengths.

```{r}
sapply(select(st_level, avg_dt, avg_dd, abs_growth_len, abs_growth_biov), function(x){sum(is.na(x))})
```

Increases in length are tightly correlated with increases in width, so using length vs biovolume is likely to yield similar results.

```{r}
ggplot(st_level, aes(x = abs_growth_len, y = abs_growth_biov)) +
  scale_x_log10() + 
  scale_y_log10() +
  geom_point(alpha = 0.2)
  #geom_smooth(se = F) 
```

This plot simply shows stage end size as a function of starting size. The lower right is empty because parasites do not shrink. The scatter indicates that growth, in absolute terms, varies quite a lot. For worms entering hosts at a small size (left side), parasites tend to grow more in definitive than intermediate hosts. For large initial sizes (right side), worms grow less, probably because there are limits on how large worms can be.

```{r}
ggplot(st_level, aes(x = stage_biov_start, y = stage_biov_end, color = Def.int)) +
  scale_x_log10() + 
  scale_y_log10() +
  geom_point(alpha = 0.2) +
  #geom_smooth(se = F) 
  labs(x = "Starting size", y = "End size", title = "Stage level")
```

This upper limit on growth is evident when we plot relative growth as a function of starting size. When starting small, parasites may grow many orders of magnitude (or not). But when starting big, they can only increase one or two orders of magnitude.

```{r}
ggplot(st_level, aes(x = stage_biov_start, y = rel_growth_biov, color = Def.int)) +
  scale_x_log10() + 
  #scale_y_log10() +
  geom_point(alpha = 0.2) +
  #geom_smooth(se = F) 
  labs(x = "Starting size", y = "Relative growth (orders of magnitude)", title = "Stage level")
```

The amount of growth a worm does at a given stage depends on how long they spend growing - more time, more growth. We can see that when we plot absolute growth as a function of development time. From this plot, it appears that, for a given length of time, worms grow more in definitive hosts.

```{r}
ggplot(st_level, aes(x = avg_dt, y = abs_growth_biov, color = Def.int)) +
  scale_x_log10() + scale_y_log10() +
  geom_point(alpha = 0.2) +
  geom_smooth(se = F) +
  labs(x = "Devo time (days)", y = "Absolute growth", title = "Stage level")
```
But that is mainly because they tend to enter the definitive host at a larger size. Large initial sizes facilitate absolute growth, because a doubling in size will result in much more mass when a worm is initially big. When we express growth in relative terms (orders of magnitude increase), then the growth rate appears more comparable in definitive and intermediate hosts.

```{r}
ggplot(st_level, aes(x = avg_dt, y = rel_growth_biov, color = Def.int)) +
  scale_x_log10() + 
  # scale_y_log10() +
  geom_point(alpha = 0.2) +
  geom_smooth(se = F, method = lm) +
  labs(x = "Devo time (days)", y = "Relative growth (orders of magnitude)", title = "Stage level")

```

There is a lot of variation in the previous plot, and one reason for this is that development is highly temperature dependent. We can try to remove temperature effects by converting development times to degree days. When we make the same plot using degree days, the relationship between growth and time is clearer. Henceforth, I use degree days as my measure of development time.

```{r}
ggplot(st_level, aes(x = avg_dd, y = rel_growth_biov, color = Def.int)) +
  scale_x_log10() + 
  # scale_y_log10() +
  geom_point(alpha = 0.2) +
  geom_smooth(se = F, method = lm) +
  labs(x = "Devo time (degree days)", y = "Relative growth (orders of magnitude)", title = "Stage level")

```


## Correlations between specificity and life history traits

Next, we look at the relationship between generalism and life history traits. There seems to be a slight negative association between the number of host records and relative growth, but the correlation is not clear.

```{r}
ggplot(st_level, aes(y = num_hosts_suspicious_removed, x = rel_growth_biov, color = Def.int)) +
  # scale_x_log10() + 
  scale_y_log10() +
  geom_point(alpha = 0.2) +
  geom_smooth(se = F, method = lm) +
  labs(y = "Host records", x = "Relative growth", title = "Stage level")
```

A similar pattern is seen for taxonomic dissimilarity.

```{r}
ggplot(st_level, aes(y = hsi_lcdb_suspcious_rem, x = rel_growth_biov, color = Def.int)) +
  # scale_x_log10() +
  # scale_y_log10() +
  geom_point(alpha = 0.2) +
  geom_smooth(method = 'lm', se = F) +
  labs(y = "Tax dissimilarity index", x = "Relative growth", title = "Stage level")
```

There are a number of species in the database that infect paratenic hosts. In paratenic hosts, it is usually assumed that the parasite undergoes little to no growth. Consequently, their size in paratenic hosts is often not reported - it is implicitly assumed that little to no growth occurs. Indeed, over 50% of paratenic stages in the database lack sizes compared to only 20% of obligate stages.

```{r}
group_by(st_level, Facultative)%>%
  summarize(total = n(), missing_size =  sum(is.na(stage_biov_end)), missing_growth = sum(is.na(rel_growth_biov)))%>%
  mutate(prop_missing_size = missing_size/total, prop_missing_growth = missing_growth/total)
```
In those cases where size was reported, growth in paratenic hosts is lower than in obligate hosts.

```{r}
ggplot(st_level, aes(x = Facultative, y = rel_growth_biov)) + 
  geom_boxplot(outlier.color = NA) +
  geom_point(position = 'jitter', alpha = 0.2)
```

Therefore, for any paratenic host stage without growth data, I set the growth to zero and then replotted the data. This makes the negative relationship clearer, especially for worms in intermediate hosts (since those are the paratenic hosts).

```{r}
st_level <- mutate(st_level, rel_growth_paratenic = if_else(Facultative == "paratenic" & is.na(rel_growth_biov), 0, rel_growth_biov)) # if paratenic, then dt = 0
```
```{r}
ggplot(st_level, aes(y = hsi_lcdb_suspcious_rem, x = rel_growth_paratenic, color = Def.int)) +
  # scale_x_log10() +
  # scale_y_log10() +
  geom_point(aes(shape = Facultative), alpha = 0.2) +
  geom_smooth(se = F, method = 'lm') +
  labs(y = "Tax dissimilarity index", x = "Relative growth, paratenics added", title = "Stage level")
```

It looks like specificity and growth are linked. Parasites grow less when they are more generalist at a given stage. What about developmental time? Is there a tradeoff between generalism and development?

It does not look like it. Stages with long development times are as generalist as those with short development times.

```{r}
ggplot(st_level, aes(y = num_hosts_suspicious_removed, x = avg_dd, color = Def.int)) +
  scale_x_log10() + scale_y_log10() +
  geom_point(alpha = 0.2) +
  geom_smooth(se = F, method = lm) +
  labs(y = "Host records", x = "Devo time, degree days", title = "Stage level")
```

```{r}
ggplot(st_level, aes(y = hsi_lcdb_suspcious_rem, x = avg_dd, color = Def.int)) +
  scale_x_log10() + 
  # scale_y_log10() +
  geom_point(alpha = 0.2) +
  geom_smooth(se = F, method = lm) +
  labs(y = "Tax dissim index", x = "Devo time, degree days", title = "Stage level")
```

In paratenic hosts, parasites are assumed to undergo no development. This was implicitly assumed in many papers used to construct the life cycle database. Let's assume that in paratenic hosts, parasites require a day at room temperature to become infective to the next host. When we replot the data, we see that these species in paratenic hosts (left side) tend to have higher generalism.

```{r}
st_level <- mutate(st_level, avg_dd_paratenic = if_else(Facultative == "paratenic" & is.na(avg_dd), 20, avg_dd)) # if paratenic, then dt = 0
```
```{r}
ggplot(st_level, aes(y = hsi_lcdb_suspcious_rem, x = avg_dd_paratenic, color = Def.int)) +
  scale_x_log10() + 
  #scale_y_log10() +
  geom_point(aes(shape = Facultative), alpha = 0.2) +
  geom_smooth(se = F) +
  labs(y = "Tax dissim index", x = "Devo time, degree days\nparatenic added", title = "Stage level")
```

Here's another way to look at it. Parasites are generalists when infecting paratenic intermediate hosts.

```{r}
ggplot(st_level, aes(x = Facultative, y = hsi_lcdb_suspcious_rem, color = Def.int)) + 
  geom_boxplot(outlier.color = NA) +
  geom_point(alpha = 0.1, position = position_jitterdodge(jitter.width = 0.3, jitter.height = 0)) +
  labs(x = "Facultative host?", y = "Taxonomic dissimilarity index")
```

Instead of looking at growth and development separately, we can also look at them together, namely as a growth rate (relative size increase per unit time). There is not a clear relationship between growth rate and host generalism, suggesting generalists do not grow slower than specialists (though they grow less).

```{r}
st_level <- mutate(st_level, rel_growth_rate_dt = rel_growth_biov/avg_dt, rel_growth_rate_dd = rel_growth_biov/avg_dd) # make growth rate variable
```
```{r}
ggplot(filter(st_level, Facultative != 'postcyclic'),
       aes(x = rel_growth_rate_dd, y = hsi_lcdb_suspcious_rem, color = Def.int)) + 
  geom_point(alpha = 0.5) +
  geom_smooth(method = lm, se = F) +
  scale_y_continuous(limits = c(1,6), breaks = c(1:6), labels = c("species", tax.ranks)) +
  labs(x = "Relative growth rate in degree days", y = "Taxonomic dissimilarity") +
  theme(legend.title = element_blank())
# ggsave(filename = "tax_dissim_vs_growth_rate.png", device = 'png')
```


## Break variables by life cycle stage

We can break this down further by looking at how our variables of interest vary across life cycles of different lengths. This plot shows how the middle hosts in long life cycles tend to have higher generalism.

```{r}
ggplot(filter(st_level, Facultative != 'postcyclic'),
       aes(x = factor(Host.no), y = hsi_lcdb_suspcious_rem, color = Def.int)) + 
  geom_boxplot(outlier.color = NA) +
  geom_point(alpha = 0.1, position = position_jitterdodge(jitter.width = 0.3, jitter.height = 0)) +
  scale_y_continuous(limits = c(1,6), breaks = c(1:6), labels = c("species", tax.ranks)) +
  labs(x = "Stage (host) in life cycle", y = "Taxonomic dissimilarity") +
  facet_wrap(~lcl_max_fac, nrow = 2)+ 
  theme(legend.title = element_blank())
```

Here are the development times in degree days across the life cycle. Development times are often higher in definitive hosts, which is probably because many of these hosts are endotherms (high temperature, many degree days). 

```{r}
ggplot(st_level, aes(x = factor(Host.no), y = avg_dd, color = Def.int)) + 
  geom_boxplot(outlier.color = NA) +
  geom_point(alpha = 0.1, position = position_jitterdodge(jitter.width = 0.3, jitter.height = 0)) +
  scale_y_log10() +
  labs(x = "Stage (host) in life cycle", y = "Devo time, degree days") +
  facet_wrap(~lcl_max_fac, nrow = 2)+ 
  theme(legend.title = element_blank())
```

Note above that the middle hosts (second, third intermediate hosts) in long life cycles exhibit long development. This pattern can be changed dramatically by assuming that worms essentially undergo no development in paratenic hosts. Under this assumption, we see that second and third intermediate hosts, those most likely to be paratenic hosts, actually have the shortest developmental times.

```{r}
ggplot(st_level, aes(x = factor(Host.no), y = avg_dd_paratenic, color = Def.int)) + 
  geom_boxplot(outlier.color = NA) +
  geom_point(alpha = 0.1, position = position_jitterdodge(jitter.width = 0.3, jitter.height = 0)) +
  scale_y_log10() +
  labs(x = "Stage (host) in life cycle", y = "Devo time, degree days") +
  facet_wrap(~lcl_max_fac, nrow = 2)+ 
  theme(legend.title = element_blank())
```

Let's next look at how growth, in absolute terms, varies across life cycles. It is usually highest in definitive hosts, again because worms tend to enter definitive hosts large, double in size a few times, and then stop growing.

```{r}
ggplot(st_level, aes(x = factor(Host.no), y = abs_growth_biov, color = Def.int)) + 
  geom_boxplot(outlier.color = NA) +
  geom_point(alpha = 0.1, position = position_jitterdodge(jitter.width = 0.3, jitter.height = 0)) +
  scale_y_log10() +
  labs(x = "Stage (host) in life cycle", y = "Absolute growth, length") +
  facet_wrap(~lcl_max_fac, nrow = 2)+ 
  theme(legend.title = element_blank())
```

When we look at relative growth, it is rather comparable across stages.

```{r}
ggplot(st_level, aes(x = factor(Host.no), y = rel_growth_biov, color = Def.int)) + 
  geom_boxplot(outlier.color = NA) +
  geom_point(alpha = 0.1, position = position_jitterdodge(jitter.width = 0.3, jitter.height = 0)) +
  # scale_y_log10() +
  labs(x = "Stage (host) in life cycle", y = "Rel growth, biovolume") +
  facet_wrap(~lcl_max_fac, nrow = 2)+ 
  theme(legend.title = element_blank())

```

However, if we assume that worms do not grow in paratenic hosts, we see that growth is especially low in second and third intermediate hosts.

```{r}
ggplot(st_level, aes(x = factor(Host.no), y = rel_growth_paratenic, color = Def.int)) + 
  geom_boxplot(outlier.color = NA) +
  geom_point(alpha = 0.1, position = position_jitterdodge(jitter.width = 0.3, jitter.height = 0)) +
  # scale_y_log10() +
  labs(x = "Stage (host) in life cycle", y = "Rel growth, length") +
  facet_wrap(~lcl_max_fac, nrow = 2)+ 
  theme(legend.title = element_blank())

```


```{r}
st_level <- mutate(st_level, Host_no_fac = if_else(Host.no > 4, as.integer(4), Host.no))%>%
  mutate(Host_no_fac = factor(Host_no_fac))
```

# Within-species correlations

```{r}
# detach(package:lme4, unload = T)
```

Up to this point, I have looked for general correlations across species while controlling for 'species effects'. However, it may be that species that are generalist at one stage of the life cycle are less so at another stage. Let's look at the correlations within species.

#### Generalism

```{r}
# select relevant columns (already species avgs) and spread the data
st_spread_hsi <- filter(st_level, stage_missing_info == 'complete')%>%
  select(Parasite.species, Host.no, hsi_lcdb_suspcious_rem, study_effort, lcl_max, lcl_min, lcl_max_fac)%>%
  spread(key = Host.no, value = hsi_lcdb_suspcious_rem)%>%
  rename(first_host = `1`, second_host = `2`, third_host = `3`, fourth_host = `4`, fifth_host = `5`)
```

In complex life cycle worms, generalism in the first and second host is uncorrelated (if anything it is positively correlated), which is not consistent with a tradeoff.

```{r}
ggplot(filter(st_spread_hsi, lcl_max > 1),
       aes(x = first_host, y = second_host)) +
  geom_point(aes(size = study_effort), alpha = 0.3) + 
  geom_smooth(se = F) +
  labs(x = "Tax dissimilarity in first host", y = "Tax dissimilarity in second host", title = "Generalism between hosts, panels are life cycle length") +
  facet_wrap(~lcl_max_fac, nrow=2)
```

```{r}
# x <- glm(first_host ~ second_host + log(study_effort+1), data = filter(st_spread_hsi, lcl_max == 2))
# summary(x)
```

#### Growth
We can also look at growth in consecutive hosts.

```{r}
# select relevant columns (already species avgs) and spread the data
st_spread_hsi <- filter(st_level, stage_missing_info == 'complete')%>%
  select(Parasite.species, Host.no, abs_growth_biov, study_effort, lcl_max, lcl_min, lcl_max_fac)%>%
  spread(key = Host.no, value = abs_growth_biov)%>%
  rename(first_host = `1`, second_host = `2`, third_host = `3`, fourth_host = `4`, fifth_host = `5`)
```

In absolute terms, worms that grow larger in their first host tend to also grow large in the second host. Presumably, size is just compounding itself across hosts - large worms can more easily accumulate additional mass in absolute terms.

```{r}
ggplot(filter(st_spread_hsi, lcl_max > 1),
       aes(x = first_host, y = second_host)) +
  geom_point(alpha = 0.3) + 
  geom_smooth(se = F, method = 'lm') +
  scale_y_log10() + scale_x_log10() +
  labs(x = "Absolute growth in first host", y = "Absolute growth in second host", title = "Growth between hosts, panels are life cycle length") +
  facet_wrap(~lcl_max_fac, nrow=2)
```

However, it relative terms, more growth in the first host leaves less opportunity for large increases in growth in the next host - there is a negative relationship between relative growth across hosts.

```{r}
# select relevant columns (already species avgs) and spread the data
st_spread_hsi <- filter(st_level, stage_missing_info == 'complete')%>%
  select(Parasite.species, Host.no, rel_growth_paratenic, study_effort, lcl_max, lcl_min, lcl_max_fac)%>%
  spread(key = Host.no, value = rel_growth_paratenic)%>%
  rename(first_host = `1`, second_host = `2`, third_host = `3`, fourth_host = `4`, fifth_host = `5`)
```
```{r}
ggplot(filter(st_spread_hsi, lcl_max > 1),
       aes(x = first_host, y = second_host)) +
  geom_point(alpha = 0.3) + 
  geom_smooth(se = F) +
  labs(x = "Relative growth in first host", y = "Relative growth in second host", title = "Growth between hosts, panels are life cycle length") +
  facet_wrap(~lcl_max_fac, nrow=2)
```

#### Devo times

There are not many examples of development times being recorded across multiple stages of a parasite life cycle. Still, no pattern is apparent.

```{r}
# select relevant columns (already species avgs) and spread the data
st_spread_hsi <- filter(st_level, stage_missing_info == 'complete')%>%
  select(Parasite.species, Host.no, avg_dd_paratenic, study_effort, lcl_max, lcl_min, lcl_max_fac)%>%
  spread(key = Host.no, value = avg_dd_paratenic)%>%
  rename(first_host = `1`, second_host = `2`, third_host = `3`, fourth_host = `4`, fifth_host = `5`)
```

```{r}
ggplot(filter(st_spread_hsi, lcl_max > 1),
       aes(x = first_host, y = second_host)) +
  geom_point(alpha = 0.3) + 
  geom_smooth(se = F, method = "lm") +
  scale_y_log10() + scale_x_log10() +
  labs(x = "Development time in first host", y = "Development time in second host", title = "Devo time, panels are life cycle length") +
  facet_wrap(~lcl_max_fac, nrow=2)
```

